---
title: "Study of Public Health and Economic Problems Caused by Severe Weather Events in the US"
output: 
    html_document:
        keep_md: true
---


## 1. Synopsis
Storms and other severe weather events can cause both public health and economic problems for communities and municipalities. In this report, we summarize the storm data from NOAA Storm Database. Our data analyis suggested that tornado has the most severe impact on public health, while flood and drought lead to highest property damage and crop damage, respectivily.


## 2. Loading and Processing the Data
The data was obtained from U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. First, we download the data from the web, and read the raw data.

```{r cache = TRUE}
filename <- "repdata%2Fdata%2FStormData.csv.bz2"
fileURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
if (!file.exists(filename)) {
    download.file(fileURL, filename)
}
raw <- read.csv(filename)
```

After reading the raw data we check the dimension of the data.

```{r}
dim(raw)
```

```{r echo = FALSE, warning = FALSE, message = FALSE}
# load all packages for data processing and plotting
library(plyr)
library(dplyr)
library(ggplot2)
library(grid)
library(gridExtra)
```

In order to see which event has the greatest effect on public health, we select the columns of interest, and summarize fatal and injure data with respect to event types.
```{r}
data <- select(raw, EVTYPE, FATALITIES : CROPDMGEXP)
harm <- ddply(data, .(EVTYPE), summarize, FATALITIES = sum(FATALITIES), INJURIES = sum(INJURIES))
fatal_ordered <- arrange(harm, desc(FATALITIES))
injure_ordered <- arrange(harm, desc(INJURIES))
```

The second thing we need to focus is that which event has the greatest economic consequence. To achieve that, we need to calculate the damage cost resulting from each event type.

```{r}
# this function is to determine exponential value
exp <- function(x) {
    if (x %in% c("h", "H")) {return(2)}
    else if (x %in% c("k", "K")) {return(3)}
    else if (x %in% c("m", "M")) {return(6)}
    else if (x %in% c("b", "B")) {return(9)}
    else if (x %in% c("", "-", "?", "+")) {return(0)}
    else {return(as.numeric(x))}    
}
```

With the funcion in hand, we can go ahead to create a new data frame that contains information needed to answer this question.

```{r cache = TRUE}
# calculate the total damage for each observation
propdmgexp <- with(data, sapply(as.character(PROPDMGEXP), exp))
propdmg <- data$PROPDMG * (10^propdmgexp)
cropdmgexp <- with(data, sapply(as.character(CROPDMGEXP), exp))
cropdmg <- data$CROPDMG * (10^cropdmgexp)

# create a new data frame that summarize the damage for each type of event
damagedata <- data %>% select(EVTYPE) %>% mutate(propdmg = propdmg, cropdmg = cropdmg)
damage <- ddply(damagedata, .(EVTYPE), summarize, propdmg = sum(propdmg), cropdmg = sum(cropdmg))

# order the prop damage and crop damage with respect to event types
prop_ordered <- arrange(damage, desc(propdmg))
crop_ordered <- arrange(damage, desc(cropdmg))
```

## 3. Results

### 3.1 Events that are most harmful to population health

Based on the data, the top-5 events that cause the most fatalities are:

```{r}
head(fatal_ordered[, c(1, 2)], n = 5)
```

The top-5 events that cause the most injuries are:

```{r}
head(injure_ordered[, c(1, 3)], n = 5)
```

To make it more clear, we use the figure below to demostrate our findings.

```{r}
p1 <- ggplot(data = fatal_ordered[1:5, ], aes(x = reorder(EVTYPE, -FATALITIES), y = FATALITIES)) + geom_bar(fill = "red", width = 0.5, stat = "identity") + xlab("Event Type") + ylab("Fatalities") + theme(legend.position="none")
p2 <- ggplot(data = injure_ordered[1:5, ], aes(x = reorder(EVTYPE, -INJURIES), y = INJURIES)) + geom_bar(fill = "red", width = 0.5, stat = "identity") + xlab("Event Type") + ylab("Injuries") + theme(legend.position="none")
grid.arrange(p1, p2, nrow = 2)
```

As the figure shows, tornado is the most harmful to population health.

### 3.2 Events that have the greatest economic consequence

Based on the data, the top-5 events that cause the most property damages are:

```{r}
head(prop_ordered[, c(1, 2)], n = 5)
```

The top-5 events that cause the most crop damages are:

```{r}
head(crop_ordered[, c(1, 3)])
```

Again we plot a figure to make it more clear.

```{r}
p3 <- ggplot(data = prop_ordered[1:5, ], aes(x = reorder(EVTYPE, -propdmg), y = propdmg)) + geom_bar(fill = "red", width = 0.5, stat = "identity") + xlab("Event Type") + ylab("Property Damage") + theme(legend.position="none")
p4 <- ggplot(data = crop_ordered[1:5, ], aes(x = reorder(EVTYPE, -cropdmg), y = cropdmg)) + geom_bar(fill = "red", width = 0.5, stat = "identity") + xlab("Event Type") + ylab("Crop Damage") + theme(legend.position="none")    
grid.arrange(p3, p4, nrow = 2)
```

Therefore, as the result shows, flood causes the most property damage, and drought cause the highest crop damage.